---
title: "Reproducible Research"
author: "Chiranthi"
date: "7/1/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Peer-graded Assignment: Course Project 1

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Loading Packages 

```{r pressure, echo=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
```

## lOading Data
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl ,destfile = "data.zip", method = "curl")
dateDownloaded <- date()
unzip(data, list = TRUE)

activitydata <- read.csv(unzip(data, "activity.csv"))
```

## Getting some insights about data
```{r}
head(activitydata)
glimpse(activitydata)
str(activitydata)
```

## converting date column to date data type
```{r}
activitydata$Date <- as.Date(paste0(as.character(activitydata$date), format = "ymd"))

```

## Histogram of the total number of steps taken each day
```{r}
Steps <- activitydata %>%
  group_by(Date) %>%
  summarise( Total = sum(steps), na.rm = TRUE)

# Drawing the histogram 
ggplot(Steps)+
  geom_bar(aes(Date, Total))
```


